### 2017-05-30 resubmission-pkg.tex
Revised submission

\documentclass{article}[11pt,oneside]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INPUTS FOR THE NEW PAPER
\newcommand{\subtitle}{Unsupervised Acquisition of Comprehensive Multiword Lexicons using Competition in an $n$-gram Lattice}
\title{Resubmission of TACL \#1137, \\ \subtitle.   \\}

\newcommand{\resubmission}{tacl.pdf}  %% revised version
\newcommand{\origdecision}{original_letter.pdf}  %% name of anonymized original-decision-letter file
\newcommand{\coverfile}{reviewer_response.pdf}  %% name of anonymized cover-letter file

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{hyperref}
\hypersetup{
    colorlinks=true, %set true if you want colored links
    linktoc=all,     %set to all if you want both sections and subsections linked
    linkcolor=blue,  %choose some color if you want links to stand out
}
\usepackage{ifthen}
\usepackage{xspace}

\usepackage{times}
\usepackage{fullpage}
\usepackage{pdfpages}

\special{papersize=8.5in,11in}
\pdfpageheight\paperheight
\pdfpagewidth\paperwidth
\setlength\topmargin{-7mm} \setlength\oddsidemargin{-0cm}
\setlength\textheight{22cm} \setlength\textwidth{15.8cm}
\setlength\columnsep{0.25in}  \newlength\titlebox \setlength\titlebox{2.00in}
\setlength\headheight{5pt}   \setlength\headsep{0pt}
\setlength\footskip{1.9cm}
\setlength\leftmargin{0.0in}
\pagestyle{plain}

\setlength{\footskip}{55pt}

\begin{document}
\maketitle

\tableofcontents

\section{Author(s) cover letter responding to the original reviews} Starts on next page.
\includepdf[pages=-,offset=0mm 6mm]{\coverfile}
\section{Revised submission} Starts on next page.
\includepdf[pages=-,offset=0mm 6mm]{\resubmission}
\section{Original decision letter and reviews} Starts on next page.
\includepdf[pages=-,offset=0mm 6mm]{\origdecision}

\end{document}